Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-05-21 manuscript/manuscript.Rmd
updates ms with error table

Results
-------
### Vital rate statistical models

### Model validation

### Climate change simulations

@@

Tables
------
```{r table_1,results='asis',message=FALSE, comment=NA, echo=FALSE}
# Make a table for average residual from each model and year effects structure
ipm_onestep <- readRDS("../montana/ipm/simulations/one_step_results/ipm_one-step_forecasts_combined.rds")
resD <- readRDS("../montana/quadBM/simulations/oneStepResids_YearEffects.rds")
resDny <- readRDS("../montana/quadBM/simulations/oneStepResids_NoYearEffects.rds")
resD$simtype <- rep("full", nrow(resD))
resDny$simtype <- rep("noyear", nrow(resDny))
resDall <- rbind(resD, resDny)
noZ <- which(resDall$lagCover!=0)
resDall <- resDall[noZ,]

stats <- ddply(ipm_onestep, .(species, year), summarise,
               mean_resid = mean(abs(resid)),
               sd_resid = sd(resid),
               min_resid = min(resid),
               max_resid = max(resid))
stats$model <- rep("IPM", nrow(stats))
colnames(stats)[2] <- "simtype"
stats2 <- ddply(resDall, .(species, simtype), summarise,
               mean_resid = mean(abs(resids)),
               sd_resid = sd(resids),
               min_resid = min(resids),
               max_resid = max(resids))
stats2$model <- rep("QBM", nrow(stats2))
stats$simtype <- stats2$simtype
stats <- rbind(stats, stats2)
stats <- stats[,c(ncol(stats),1:(ncol(stats)-1))]
stats$simtype <- rep(c("Full", "Climate"),(nrow(stats)/2))
colnames(stats) <- c("Model", "Species", "Year Structure", "Mean", "SD", "Min", "Max")
print.xtable(xtable(stats, caption = "Table of residual summary statistics."),type="latex", comment=FALSE, include.rownames=FALSE)
```

@@

```{r figure_2, dependson="plot-options", fig.cap="One step forecasts from IPM version.",fig.width=8, fig.height=8, cache=FALSE, results='hide'}
resD <- readRDS("../montana/quadBM/simulations/oneStepResids_YearEffects.rds")
resDny <- readRDS("../montana/quadBM/simulations/oneStepResids_NoYearEffects.rds")
resD$simtype <- rep("full", nrow(resD))
resDny$simtype <- rep("noyear", nrow(resDny))
resDall <- rbind(resD, resDny)
noZ <- which(resDall$lagCover!=0)
resDall <- resDall[noZ,]

myCols <- c("#237DA4", "#7ECAD0")
# ggplot(resDall, aes(x=resids, fill=simtype))+
#   geom_histogram(alpha=0.5)+
#   facet_grid(.~species, scales = "free_x")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   scale_fill_manual(values=myCols)

resAgg <- ddply(resDall, .(year, species, simtype), summarise,
                avg_prediction = mean(cover)*100,
                up_prediction = quantile(cover, probs = 0.875)*100,
                down_prediction = quantile(cover, probs = 0.125)*100,
                avg_obs = mean(propCover)*100)
resAgg$model <- rep("QBM", nrow(resAgg))

ipm_onestep <- readRDS("../montana/ipm/simulations/one_step_results/ipm_one-step_forecasts_combined.rds")
ipm_onestep$t1 <- ipm_onestep$t1+1900

resAggipm <- ddply(ipm_onestep, .(t1, species, year), summarise,
                avg_prediction = mean(cover.t1)*100,
                up_prediction = quantile(cover.t1, probs = 0.875)*100,
                down_prediction = quantile(cover.t1, probs = 0.125)*100,
                avg_obs = mean(obs.cover.t1)*100)
colnames(resAggipm)[1:3] <- c("year", "species", "simtype")
resAggipm$simtype[which(resAggipm$simtype=="ayear effect")] <- "full"
resAggipm$simtype[which(resAggipm$simtype!="full")] <- "noyear"
resAggipm$model <- rep("IPM", nrow(resAggipm))

resAgg_both <- rbind(resAgg, resAggipm)

ggplot(subset(resAgg_both))+
  geom_ribbon(aes(x=year, ymin=down_prediction, ymax=up_prediction, fill=simtype), alpha=0.3)+
  geom_line(aes(x=year, y=avg_prediction, color=simtype), size=1.5)+
  geom_point(aes(x=year, y=avg_obs), shape=1, size=3)+
  facet_grid(species~model, scales="free")+
  scale_fill_manual(values=c(myCols[1], "darkorange"), labels=c("Year effects", "No year effects"))+
  scale_color_manual(values=c(myCols[1], "darkorange"))+
  guides(color=FALSE)

ggplot(subset(resAgg_both))+
  geom_ribbon(aes(x=year, ymin=down_prediction, ymax=up_prediction, fill=model), alpha=0.3)+
  geom_line(aes(x=year, y=avg_prediction, color=model), size=1.5)+
  geom_point(aes(x=year, y=avg_obs), shape=1, size=3)+
  facet_grid(species~simtype, scales="free")+
  scale_fill_manual(values=c(myCols[1], "darkorange"))+
  scale_color_manual(values=c(myCols[1], "darkorange"))+
  guides(color=FALSE)

@@

```{r figure_3, dependson="plot-options", fig.cap="One step forecasts from IPM version.",fig.width=8, fig.height=8, cache=FALSE, results='hide', eval=FALSE}
ipm_onestep2 <- subset(ipm_onestep, year=="ayear effect")
resDall2 <- subset(resDall, simtype=="full")
ipm_qbm_onestep <- data.frame(year = c(ipm_onestep2$t1, resDall2$year),
                              resids = c(ipm_onestep2$resid, resDall2$resids),
                              model = c(rep("IPM", nrow(ipm_onestep2)),
                                        rep("QBM", nrow(resDall2))),
                              species = c(ipm_onestep2$species, resDall2$species))
g1 <- ggplot(subset(ipm_qbm_onestep, species=="BOGR"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  coord_cartesian(ylim = c(-50, 50))+
  ggtitle(expression(italic("B. gracilis")))+
  guides(fill=FALSE)
g2 <- ggplot(subset(ipm_qbm_onestep, species=="HECO"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  coord_cartesian(ylim = c(-10, 5))+
  ggtitle(expression(italic("H. comata")))+
  guides(fill=guide_legend(title=""))+
  theme(legend.position=c(0.5,0.2))
g3 <- ggplot(subset(ipm_qbm_onestep, species=="PASM"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("P. smithii")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-1.5, 1.5))
g4 <- ggplot(subset(ipm_qbm_onestep, species=="POSE"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("Poa secunda")))+
  coord_cartesian(ylim = c(-10, 15))+
  guides(fill=FALSE)
pFinal <- grid.arrange(g1,g2,g3,g4, ncol=2, nrow=2)
pFinal
```

>>>>

Results
-------
### Comparison of Forecast Models

### Forecasting Climate Change Impacts

@@

Tables
------
```{r table_1,results='asis',message=FALSE, comment=NA, echo=FALSE}
# Make a table for average residual from each model and year effects structure
ipm_onestep <- readRDS("../montana/ipm/simulations/one_step_results/ipm_one-step_forecasts_combined.rds")
qbm_onestep <- readRDS("../montana/quadBM/simulations/validation_results/qbm_one-step_forecasts_combined.rds")

#IPM
tmpipm <- ddply(ipm_onestep, .(quad, t1, species), summarise,
                mean_pred = mean(cover.t1),
                var_pred = var(cover.t1),
                mean_obs = mean(obs.cover.t1),
                resid = mean(resid))
ipmstats <- ddply(tmpipm, .(species), summarise,
                rho = cor(mean_pred, mean_obs),
                mae = mean(abs(resid)),
                var = mean(var_pred),
                mean_cover = mean(mean_obs*100))
ipmstats$model <- "IPM"

#QBM
tmpqbm <- ddply(qbm_onestep, .(quad, year, species), summarise,
                mean_pred = mean(predcov),
                var_pred = var(predcov),
                mean_obs = mean(obscov),
                resid = mean(resid))
qbmstats <- ddply(tmpqbm, .(species), summarise,
                rho = cor(mean_pred, mean_obs),
                mae = mean(abs(resid)),
                var = mean(var_pred),
                mean_cover = mean(mean_obs*100))
qbmstats$model <- "QBM"

stats_table <- rbind(ipmstats, qbmstats)
stats_table <- stats_table[ order(stats_table[,"species"]), ]
stats_table <- stats_table[,c("model", "species", "mae",
                              "rho","mean_cover")]

colnames(stats_table) <- c("Model", "Species", "MAE", "Accuracy", "Mean Obs. Cover")
print.xtable(xtable(stats_table, caption = "Mean absolute error (MAE) and accuracy (Pearson's $\\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation"),type="latex", comment=FALSE, include.rownames=FALSE)
```

@@

```{r figure_2, dependson="plot-options", fig.cap="One step forecasts from IPM version.",fig.width=8, fig.height=8, cache=FALSE, results='hide'}
# resD <- readRDS("../montana/quadBM/simulations/oneStepResids_YearEffects.rds")
# resDny <- readRDS("../montana/quadBM/simulations/oneStepResids_NoYearEffects.rds")
# resD$simtype <- rep("full", nrow(resD))
# resDny$simtype <- rep("noyear", nrow(resDny))
# resDall <- rbind(resD, resDny)
# noZ <- which(resDall$lagCover!=0)
# resDall <- resDall[noZ,]
# 
# myCols <- c("#237DA4", "#7ECAD0")
# # ggplot(resDall, aes(x=resids, fill=simtype))+
# #   geom_histogram(alpha=0.5)+
# #   facet_grid(.~species, scales = "free_x")+
# #   xlab("Year")+
# #   ylab("Residuals (% cover difference)")+
# #   theme_bw()+
# #   scale_fill_manual(values=myCols)
# 
# resAgg <- ddply(resDall, .(year, species, simtype), summarise,
#                 avg_prediction = mean(cover)*100,
#                 up_prediction = quantile(cover, probs = 0.875)*100,
#                 down_prediction = quantile(cover, probs = 0.125)*100,
#                 avg_obs = mean(propCover)*100)
# resAgg$model <- rep("QBM", nrow(resAgg))
# 
# 
# 
# ipm_onestep <- readRDS("../montana/ipm/simulations/one_step_results/ipm_one-step_forecasts_combined.rds")
# ipm_onestep$t1 <- ipm_onestep$t1+1900
# 
# resAggipm <- ddply(ipm_onestep, .(t1, species, year), summarise,
#                 avg_prediction = mean(cover.t1)*100,
#                 up_prediction = quantile(cover.t1, probs = 0.875)*100,
#                 down_prediction = quantile(cover.t1, probs = 0.125)*100,
#                 avg_obs = mean(obs.cover.t1)*100)
# colnames(resAggipm)[1:3] <- c("year", "species", "simtype")
# resAggipm$simtype[which(resAggipm$simtype=="ayear effect")] <- "full"
# resAggipm$simtype[which(resAggipm$simtype!="full")] <- "noyear"
# resAggipm$model <- rep("IPM", nrow(resAggipm))
# 
# resAgg_both <- rbind(resAgg, resAggipm)
# 
# ggplot(subset(resAgg_both))+
#   geom_ribbon(aes(x=year, ymin=down_prediction, ymax=up_prediction, fill=simtype), alpha=0.3)+
#   geom_line(aes(x=year, y=avg_prediction, color=simtype), size=1.5)+
#   geom_point(aes(x=year, y=avg_obs), shape=1, size=3)+
#   facet_grid(species~model, scales="free")+
#   scale_fill_manual(values=c(myCols[1], "darkorange"), labels=c("Year effects", "No year effects"))+
#   scale_color_manual(values=c(myCols[1], "darkorange"))+
#   guides(color=FALSE)
# 
# ggplot(subset(resAgg_both))+
#   geom_ribbon(aes(x=year, ymin=down_prediction, ymax=up_prediction, fill=model), alpha=0.3)+
#   geom_line(aes(x=year, y=avg_prediction, color=model), size=1.5)+
#   geom_point(aes(x=year, y=avg_obs), shape=1, size=3)+
#   facet_grid(species~simtype, scales="free")+
#   scale_fill_manual(values=c(myCols[1], "darkorange"))+
#   scale_color_manual(values=c(myCols[1], "darkorange"))+
#   guides(color=FALSE)

@@

```{r figure_3, dependson="plot-options", fig.cap="One step forecasts from IPM version.",fig.width=8, fig.height=8, cache=FALSE, results='hide', eval=FALSE}
# ipm_onestep2 <- subset(ipm_onestep, year=="ayear effect")
# resDall2 <- subset(resDall, simtype=="full")
# ipm_qbm_onestep <- data.frame(year = c(ipm_onestep2$t1, resDall2$year),
#                               resids = c(ipm_onestep2$resid, resDall2$resids),
#                               model = c(rep("IPM", nrow(ipm_onestep2)),
#                                         rep("QBM", nrow(resDall2))),
#                               species = c(ipm_onestep2$species, resDall2$species))
# g1 <- ggplot(subset(ipm_qbm_onestep, species=="BOGR"))+
#   geom_hline(aes(yintercept=0), linetype=2)+
#   geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
#   scale_fill_manual(values=myCols)+
#   coord_cartesian(ylim = c(-50, 50))+
#   ggtitle(expression(italic("B. gracilis")))+
#   guides(fill=FALSE)
# g2 <- ggplot(subset(ipm_qbm_onestep, species=="HECO"))+
#   geom_hline(aes(yintercept=0), linetype=2)+
#   geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
#   scale_fill_manual(values=myCols)+
#   coord_cartesian(ylim = c(-10, 5))+
#   ggtitle(expression(italic("H. comata")))+
#   guides(fill=guide_legend(title=""))+
#   theme(legend.position=c(0.5,0.2))
# g3 <- ggplot(subset(ipm_qbm_onestep, species=="PASM"))+
#   geom_hline(aes(yintercept=0), linetype=2)+
#   geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
#   scale_fill_manual(values=myCols)+
#   ggtitle(expression(italic("P. smithii")))+
#   guides(fill=FALSE)+
#   coord_cartesian(ylim = c(-1.5, 1.5))
# g4 <- ggplot(subset(ipm_qbm_onestep, species=="POSE"))+
#   geom_hline(aes(yintercept=0), linetype=2)+
#   geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
#   scale_fill_manual(values=myCols)+
#   ggtitle(expression(italic("Poa secunda")))+
#   coord_cartesian(ylim = c(-10, 15))+
#   guides(fill=FALSE)
# pFinal <- grid.arrange(g1,g2,g3,g4, ncol=2, nrow=2)
# pFinal
```

### 2015-05-21 manuscript/manuscript.tex
updates ms with error table

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit analogous models of extinction probability, percent cover
increase/decrease, and quadrat colonization for each species. We
describe the statistical models separately since fitting the models
required different approaches at the individual and quadrat levels. All
models contain four climate covariate that we chose \emph{a priori}:
fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and mean spring temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year.

@@

\paragraph{Vital rate models: individual
level}\label{vital-rate-models-individual-level}

@@

\paragraph{Vital rate models: quadrat
level}\label{vital-rate-models-quadrat-level}

The statistical approach used to model vital rates using aggregated data
depends on the type of data collected. In our case, and as is often the
case with census data, we have percent cover data (which can easily be
transformed to proportion data, of course). We first considered fitting
three vital rate models analagous to those we fit at the individual
level: one for probability of extirpation within a quadrat (analagous to
survival), one for cover change within a quadrat (analagous to growth),
and one for probability of colonization within a quadrat (analagous to
recruitment). However, within-quadrat extirpation and colonization
events were rare in our time series ($N=9$ and $N=10$, respectively).
Given the broad spatial distribution of the quadrats we are studying, it
is safe to assume that these events are in fact rare enough to be
ignored for our purposes. So we constrained our statistical modeling of
vital rates at the population level to change in percent cover within
quadrats. For most of the remaining discussion of statistical modeling
we refer to proportion data, which is simply percent data divided by
100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in quadrat-based population model the
simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change
($G$) from time $t$ to $t+1$

@@

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. \#). The log normal likelihood
includes a truncation (\$T{[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `r2jags'
(Su and Yajima 2012) to connect R to JAGS. We obtained posterior
distributions for all model parameters from three parallel MCMC chains
run for 50,000 iterations, after discarding an initial 50,000
iterations. We assessed convergence visually and using the Gelman and
Rubin (1992) diagnostic in the R package `coda' (Plummer et al. 2006).
Scale reduction factors for all parameters were less than 1.01,
indicating convergence. For the purposes of introducing stochasticity in
our population models, we saved the final 1,000 iterations from each of
the three MCMC chains for all parameters to be used as randomly drawn
values during population simulation.

@@

\subsubsection{Vital rate statistical
models}\label{vital-rate-statistical-models}

\subsubsection{Model validation}\label{model-validation-1}

\subsubsection{Climate change
simulations}\label{climate-change-simulations}

@@

\begin{table}[ht]
\centering
\begin{tabular}{lllrrrr}
  \hline
Model & Species & Year Structure & Mean & SD & Min & Max \\ 
  \hline
IPM & BOGR & Full & 3.42 & 5.78 & -46.52 & 43.05 \\ 
  IPM & BOGR & Climate & 5.77 & 11.40 & -230.10 & 47.81 \\ 
  IPM & HECO & Full & 0.56 & 0.95 & -10.99 & 4.33 \\ 
  IPM & HECO & Climate & 0.60 & 1.05 & -8.65 & 6.67 \\ 
  IPM & PASM & Full & 0.17 & 0.30 & -4.29 & 1.04 \\ 
  IPM & PASM & Climate & 0.14 & 0.20 & -2.04 & 0.99 \\ 
  IPM & POSE & Full & 0.91 & 1.65 & -30.73 & 7.34 \\ 
  IPM & POSE & Climate & 0.72 & 1.51 & -6.08 & 14.32 \\ 
  QBM & BOGR & Full & 6.80 & 12.45 & -81.48 & 52.88 \\ 
  QBM & BOGR & Climate & 7.69 & 13.61 & -94.71 & 54.33 \\ 
  QBM & HECO & Full & 0.94 & 2.28 & -68.53 & 7.84 \\ 
  QBM & HECO & Climate & 1.26 & 3.49 & -94.57 & 7.91 \\ 
  QBM & PASM & Full & 0.26 & 0.50 & -15.13 & 1.56 \\ 
  QBM & PASM & Climate & 0.27 & 0.53 & -9.69 & 1.67 \\ 
  QBM & POSE & Full & 0.86 & 2.07 & -58.34 & 13.97 \\ 
  QBM & POSE & Climate & 1.12 & 2.47 & -73.14 & 15.18 \\ 
   \hline
\end{tabular}
\caption{Table of residual summary statistics.} 
\end{table}

@@

\includegraphics{components/figure/manuscript-figure_21.pdf}
\includegraphics{components/figure/manuscript-figure_22.pdf}

@@

Su, Y., and M. Yajima. 2012. R2jags: A Package for Running jags from R.
http://CRAN. R-project. org/package= R2jags.

>>>>

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit a single regression model for population growth. We describe the
statistical models separately since fitting the models required
different approaches. All models contain five climate covariate that we
chose \emph{a priori}: ``water year'' precipitation at \emph{t}-1
(lagppt); fall through spring precipitation at \emph{t}-1 and \emph{t}-2
(ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1
and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}
is the observation year.

@@

\paragraph{Vital rate models at the individual
level}\label{vital-rate-models-at-the-individual-level}

@@

\paragraph{Population model at the quadrat
level}\label{population-model-at-the-quadrat-level}

The statistical approach used to model vital rates using aggregated data
depends on the type of data collected. In our case, and as is often the
case with census data, we have percent cover data (which can easily be
transformed to proportion data, of course). We first considered fitting
three vital rate models analagous to those we fit at the individual
level: one for probability of extirpation within a quadrat (analagous to
survival), one for cover change within a quadrat (analagous to growth),
and one for probability of colonization within a quadrat (analagous to
recruitment). However, within-quadrat extirpation and colonization
events were rare in our time series ($N=9$ and $N=10$, respectively
across all species). Given the broad spatial distribution of the
quadrats we are studying, it is safe to assume that these events are in
fact rare enough to be ignored for our purposes. So we constrained our
statistical modeling of vital rates at the population level to change in
percent cover within quadrats. For the remaining discussion of
statistical modeling we refer to proportion data, which is simply
percent data divided by 100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in a quadrat-based population model
the simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change
($G$) from time $t$ to $t+1$ is

@@

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. \#). The log normal likelihood
includes a truncation ($T${[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `rjags'
to connect R to JAGS. We obtained posterior distributions for all model
parameters from three parallel MCMC chains run for 50,000 iterations,
thinned by 50, after discarding an initial 50,000 iterations. We
assessed convergence visually and using the Gelman and Rubin (1992)
diagnostic in the R package `coda' (Plummer et al. 2006). Scale
reduction factors for all parameters were less than 1.01, indicating
convergence. For the purposes of introducing stochasticity in our
population models, we saved the final 1,000 iterations from each of the
three MCMC chains for all parameters to be used as randomly drawn values
during population simulation.

@@

\subsubsection{Comparison of Forecast
Models}\label{comparison-of-forecast-models}

\subsubsection{Forecasting Climate Change
Impacts}\label{forecasting-climate-change-impacts}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llrrr}
  \hline
Model & Species & MAE & Accuracy & Mean Obs. Cover \\ 
  \hline
IPM & BOGR & 4.77 & 0.90 & 9.16 \\ 
  QBM & BOGR & 5.30 & 0.88 & 9.22 \\ 
  IPM & HECO & 0.60 & 0.80 & 1.22 \\ 
  QBM & HECO & 2.03 & 0.84 & 1.25 \\ 
  IPM & PASM & 0.18 & 0.65 & 0.40 \\ 
  QBM & PASM & 0.17 & 0.87 & 0.40 \\ 
  IPM & POSE & 0.78 & 0.61 & 1.23 \\ 
  QBM & POSE & 1.30 & 0.57 & 1.25 \\ 
   \hline
\end{tabular}
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation} 
\end{table}

@@

@@
